---
output: html_document
---
---
title: "Reproducible Research Course Project 1"
author: "Dobs"
date: "December 26, 2016"
output: html_document

## Loading and preprocessing the data

```{r setup, include=TRUE}
library(knitr)
opts_chunk$set(echo = TRUE)
setwd("C:/Users/Una/RepData_PeerAssessment1")
unzip(zipfile="activity.zip")
activity_data <- read.csv("activity.csv")
str(activity_data)
```
## What is mean total number of steps taken per day?

```{r}
library(ggplot2)
library(plyr)
totals <- ddply(activity_data, .(date), summarise, steps_per_date = sum(steps))
summary(totals)
hist(totals$steps_per_date, main = "Summary Activity Report", col="gray", xlab="Total Num of Steps taken each day")
activity_mean <-mean(totals$steps_per_date, na.rm=TRUE)
activity_median <- median(totals$steps_per_date, na.rm=TRUE)
activity_mean
activity_median
```

## What is the average daily activity pattern?
```{r}
avg_steps_per_interval <- aggregate(steps ~ interval, activity_data, FUN = mean, na.rm = TRUE)

plot(avg_steps_per_interval$interval,avg_steps_per_interval$steps, type="l", xlab="Total 5 min Interval", ylab="Average num of steps",main="Average Daily Activity Pattern")

max_interval <- avg_steps_per_interval[which.max(avg_steps_per_interval$steps),]
max_interval
```
## Imputing missing values
```{r}
colSums_na <-colSums(is.na(activity_data))
colSums_na

get_steps_value <- function(steps, interval) {
     if (is.na(steps))
         new_steps_value <- (avg_steps_per_interval[avg_steps_per_interval$interval==interval, "steps"])
     else
         new_steps_value <- c(steps)
     return(new_steps_value)
 }

activity_data_enrich <- activity_data

activity_data_enrich$steps <- mapply(get_steps_value, activity_data_enrich$steps, activity_data_enrich$interval)


totals.enrich <- ddply(activity_data_enrich, .(date), summarise, steps_per_date = sum(steps))
totals.enrich$date_dt = as.factor(as.Date(totals.enrich$date))

head(totals.enrich)

hist(totals.enrich$steps_per_date, main = "Summary Activity Report", col="green", xlab="Total Num of Steps taken each day")
activity_enr_mean <-mean(totals.enrich$steps_per_date, na.rm=TRUE)
activity_enr_median <- median(totals.enrich$steps_per_date, na.rm=TRUE)

activity_enr_mean
activity_enr_median
```
What is the impact of imputing missing data on the estimates of the total daily number of steps?
-> the mean and the median are equal

## Are there differences in activity patterns between weekdays and weekends?
```{r}

weekdays.str <- c("Mon", "Tue", "Wed", "Thu", "Fri")

totals.enrich$weekend_ind = as.factor(ifelse(is.element(weekdays(as.Date(totals.enrich$date), abbr = TRUE ),weekdays.str), "0", "1"))

activity_data_enrich$weekend_ind = as.factor(ifelse(is.element(weekdays(as.Date(activity_data_enrich$date), abbr = TRUE ),weekdays.str), "0", "1"))

avg_steps_per_interval.enrich <- aggregate(steps ~ interval + weekend_ind, activity_data_enrich, FUN = mean, na.rm = TRUE)

avg_steps_per_interval.enrich_0 <- subset(avg_steps_per_interval.enrich, weekend_ind == '0')
avg_steps_per_interval.enrich_1 <- subset(avg_steps_per_interval.enrich, weekend_ind == '1')

par(mfrow=c(1,2))
plot(avg_steps_per_interval.enrich_0$interval,avg_steps_per_interval.enrich_0$steps, type="l", xlab="Total 5 min Interval", ylab="Average num of steps",main="Daily Activity Weekdays")

plot(avg_steps_per_interval.enrich_1$interval,avg_steps_per_interval.enrich_1$steps, type="l", xlab="Total 5 min Interval", ylab="Average num of steps",main="Daily Activity Weekends")
```


## People in average are more active on weekends than on weekdays


